---
title: "Financial Metrics of S&P 500 Companies"
execute:
  echo: FALSE
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(scales)
library(plotly)

financials <- 
read_csv("financials.csv") |>
janitor::clean_names() |>
mutate(percent_eps = earnings_share / price)

#removed outliers to better visualize eps
Q1 <- quantile(financials$percent_eps, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$percent_eps, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower <- Q1 - 1.5 * IQR
upper <- Q3 + 1.5 * IQR

#removed outliers for price_earnings
Q1 <- quantile(financials$price_earnings, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_earnings, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower1 <- Q1 - 1.5 * IQR
upper1 <- Q3 + 1.5 * IQR

#removed outliers for price_book 
Q1 <- quantile(financials$price_book, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_book, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower2 <- Q1 - 1.5 * IQR
upper2 <- Q3 + 1.5 * IQR

#removing all outliers
financials_filtered <- financials |>
  filter(percent_eps >= lower, percent_eps <= upper,
         price_earnings >= lower1, price_earnings <=              upper1, price_book >= lower2, price_book <=              upper2)

```

```{r}
#| label: market_cap_to_eps
#| message: false

f <- financials_filtered |>
ggplot(aes(x = market_cap,
           y = percent_eps,
           color = sector)) +
geom_point(alpha = 0.5, size = 0.7) +
geom_smooth(se = FALSE,
            formula = 'y ~ x',
            method = lm) +
 scale_x_continuous(labels = label_dollar(scale = 1/10^9,                              suffix = "B"),
                    breaks = seq(0, 200e12, by = 100e9)) +
 scale_y_continuous(labels = scales::percent) +
labs(title = "Market Cap vs Earnings per Share by Sector",
     subtitle = "Earnings per share does not increase with market cap or sector",
     x = "Market Capitalization in Billions of Dollars",
     y = "Percent Earnings per Share", 
     color = "Sector") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 30, hjust = 1))

ggplotly(f)
```

**Overview**

Trends in metrics used to determine stock performance vary significantly by sector. This project uses data from S&P 500 companies in July 2020. Columns relating to popular financial metrics, including earnings per share and market capitalization, were used to analyze trends throughout different sectors. Earnings per share increases with market capitalization in sectors experiencing consistent levels of demand. 

**What is the S&P 500?**

The S&P 500 is one of the most widely-used US stock market indexes. It includes 503 leading publicly traded companies in the US, and has been used to provide insight into the direction of the stock market as a whole.
