---
title: "Financial Metrics of S&P 500 Companies"
execute:
  echo: FALSE
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(scales)
library(plotly)

financials <- 
read_csv("financials.csv") |>
janitor::clean_names() |>
mutate(percent_eps = earnings_share / price)

#removed outliers to better visualize eps
Q1 <- quantile(financials$percent_eps, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$percent_eps, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower <- Q1 - 3 * IQR
upper <- Q3 + 3 * IQR

#removed outliers for price_earnings
Q1 <- quantile(financials$price_earnings, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_earnings, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower1 <- Q1 - 3 * IQR
upper1 <- Q3 + 3 * IQR

#removed outliers for price_book 
Q1 <- quantile(financials$price_book, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_book, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower2 <- Q1 - 3 * IQR
upper2 <- Q3 + 3 * IQR

#removing all outliers
financials_filtered <- financials |>
  filter(percent_eps >= lower, percent_eps <= upper,
         price_earnings >= lower1, price_earnings <=              upper1, price_book >= lower2, price_book <=              upper2)

```

```{r}
#| label: market_cap_to_eps
#| message: false

#adjusted point size from 1 to 0.7
#adjusted x axis scale to dollars in billions, changed scale from 100 to 400 to make it easier to read
#added lines of best fit to better observe trends
f <- financials_filtered |>
ggplot(aes(x = market_cap,
           y = percent_eps,
           color = sector)) +
geom_point(alpha = 0.5, size = 0.7) +
geom_smooth(se = FALSE,
            formula = 'y ~ x',
            method = lm) +
 scale_x_continuous(labels = label_dollar(scale = 1/10^9,                              suffix = "B"),
                    breaks = seq(0, 200e12, by = 100e9)) +
 scale_y_continuous(labels = scales::percent) +
labs(title = "Market Cap vs Earnings per Share by Sector",
     subtitle = "Earnings per share does not increase with market cap or sector",
     x = "Market Capitalization in Billions of Dollars",
     y = "Percent Earnings per Share", 
     color = "Sector") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 30, hjust = 1))

#code for interactive graph, easier to view trends for specific sectors
ggplotly(f)
```

**Overview**
Trends in financial metrics used to evaluate stock performance vary significantly by sector. This project uses data from S&P 500 companies in July 2020. Commonly used financial metrics, such as earnings per share and market capitalization, were used to analyze trends throughout different stock market sectors. Earnings per share increases with market capitalization in sectors experiencing high or stable demand. 

**What is the S&P 500?**

The S&P 500 is one of the most widely-used US stock market indexes. It includes 503 leading publicly traded companies in the US, and has been used to provide insight into the direction of the stock market as a whole.
