---
title: "Price-to-Earnings Ratio by Sector"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(scales)

financials <- 
read_csv("financials.csv") |>
janitor::clean_names() |>
mutate(percent_eps = earnings_share / price)

#removed outliers to better visualize eps
Q1 <- quantile(financials$percent_eps, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$percent_eps, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower <- Q1 - 3 * IQR
upper <- Q3 + 3 * IQR

#removed outliers for price_earnings
Q1 <- quantile(financials$price_earnings, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_earnings, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower1 <- Q1 - 3 * IQR
upper1 <- Q3 + 3 * IQR

#removed outliers for price_book 
Q1 <- quantile(financials$price_book, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_book, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower2 <- Q1 - 3 * IQR
upper2 <- Q3 + 3 * IQR

#removing all outliers
financials_filtered <- financials |>
  filter(percent_eps >= lower, percent_eps <= upper,
         price_earnings >= lower1, price_earnings <=              upper1, price_book >= lower2, price_book <=              upper2)

```

The **price-to-earnings** ratio is a company's current share price divided by its earnings per share. A high price-to-earnings ratio suggests that a stock's market price is high compared to its earnings. This could imply the stock is overvalued or that investors expect the stock to grow in the near future. A low price-to-earnings ratio could suggest the opposite. The energy and materials sectors, each highly sensitive to shifts in the economy and global politics, have the highest median price-to-earnings ratios. 

```{r}
#| label: sector_to_pe
#| message: false 
#shifted x axis scale to make boxplot easier to read
#reordered by highest to lowest price/earnings ratio
financials_filtered |>
  ggplot(aes(x = price_earnings,
             y = fct_reorder(sector, price_earnings)))+
  geom_boxplot(fill = "light blue", alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 60, by = 4)) +
  coord_cartesian(xlim = c(0, 60), expand = FALSE) +
  labs(title = "Price/Earnings Ratio by Sector",
       x = "Price/Earnings",
       y = NULL) +
  theme_minimal()
```



