---
title: "Price-to-Book Ratio by Sector"
execute:
  echo: false
---

The **price-to-book** ratio is a company's current share price divided by its book value, or net assets, per share. Essentially, this ratio compares a company's net worth to the value given to it by the market. A high price-to-book ratio could suggest that a stock is overvalued, while a low price-to-book ratio could suggest the opposite. Information technology has the highest median price-to-book ratio, while financials has the lowest, possibly due to the amount of assets companies in these sectors tend to have. 

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(scales)

financials <- 
read_csv("financials.csv") |>
janitor::clean_names() |>
mutate(percent_eps = earnings_share / price)

#removed outliers to better visualize eps
Q1 <- quantile(financials$percent_eps, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$percent_eps, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower <- Q1 - 3 * IQR
upper <- Q3 + 3 * IQR

#removed outliers for price_earnings
Q1 <- quantile(financials$price_earnings, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_earnings, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower1 <- Q1 - 3 * IQR
upper1 <- Q3 + 3 * IQR

#removed outliers for price_book 
Q1 <- quantile(financials$price_book, 0.25, na.rm = TRUE)
Q3 <- quantile(financials$price_book, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower2 <- Q1 - 3 * IQR
upper2 <- Q3 + 3 * IQR

#removing all outliers
financials_filtered <- financials |>
  filter(percent_eps >= lower, percent_eps <= upper,
         price_earnings >= lower1, price_earnings <=              upper1, price_book >= lower2, price_book <=              upper2)

#removing non-finite values
financials_filtered2 <- financials_filtered %>%
  filter(is.finite(ebitda) & is.finite(price_earnings) & ebitda > 0)
```


```{r}
#| label: sector to pb 
#| message: false

financials_filtered |>
  ggplot(aes(x = price_book, y = fct_reorder(sector, price_book,
                                             .na_rm = TRUE))) +
  geom_boxplot(fill = "light blue", alpha = 0.5) + scale_x_continuous(breaks = seq(0, 20, by = 2)) +
  coord_cartesian(xlim = c(0, 20), expand = FALSE) +
    labs(title = "Price/Book Ratio by Sector",
       x = "Price/Book",
       y = NULL) +
  theme_minimal()
```

